Patient-Specific Neural Field Modeling from Micro-Electrode Data

### 2010-07-26 ltx/data_paper/DataPaper.tex
added andrew to data paper
committing for merge
correction to EM algorithm
added to section on estimating support for kernel

\title{Patient-Specific Neural Field Modeling from Mirco-Electrode Data}
\author{In no specific order: Parham Aram, Dean R. Freestone, Kenneth Scerri, Michael Dewar,\\
 Jacob A. Donoghue, Sydney S. Cash, Leigh R. Hochberg,\\
 David B. Grayden and Visakan Kadirkamanathan  }

@@

During the time-course of the iEEG monitoring several seizures were recorded. This study focuses on the local field potential (LFP) data recorded with the micro-electrode during a seizure.

The first pre-processing step was to identify channels that had very poor signal quality. These channels were set to the $nan$ data type. Following this, the data was re-referenced to a common average montage to remove common mode artifact and the effect of the reference electrode. Next the data was low-pass filtered with a cut-off frequency of 100~Hz. The data was then resampled from 30~kHz to 5~kHz. To enable spatial visualisation and spatial frequency analysis, the corrupted channels, set to $nan$, were interpolated using all neighbouring channels.

@@

The spectral properties of the recorded time series for each of the time segments are shown in Figure~\ref{fig:TemporalFreqObservation}. Each plot shows an averaged (across channels) spectrum the data. The pre-seizure period shows a decay in power at a typical $1/f^2$ rate. The seizure period has higher power at low and high ends of the spectrum. In addition, there is a peak in the power of the LFPs at ~51~Hz during the seizure. This peak would be difficult to observe in European and Asian countries due to the 50~Hz mains power frequency. The post-seizure period shows a large increase in lower frequency power...

Figure~\ref{fig:SpatialFreqObservation} shows the results of spatial frequency analysis of the observed neural field. Throughout this paper the spatial frequency shall be denoted by $\nu$. Each subplot is an average (across time) spatial frequency during either the pre-seizure, seizure or post-seizure periods. The spatial frequency of the neural field is governed by disturbances (inputs from outside the field) and the underlying connectivity structure. The symmetry in the plots provides some evidence for homogeneous and isotropic connectivity.

Figure~\ref{fig:DiagSpatialFreqObservation} shows diagonal cross-sections of the spatial frequency plots. From this figure we can see that the spatial cut-off frequency (-3~dB point) is $\nu_c \approx 0.26$~Hz. The profile of the diagonal cross-section indicates that the field is adequately sampled from the micro-electrode array, where the power drops by approximately 25~dB at the maximum observable spatial frequency (according to Shannon's sampling theorem) indicating spatial aliasing will have an insignificant effect. The spatial cut-off frequency provides information on the complexity of the model required to represent the neural field. Considering the field is band-limited, we can represent the continuous field by a finite set of continuous basis functions. According the Shannon's sampling theorem, the minimum spacing between basis functions, $\Delta_{\phi}$, to describe the field is
\begin{equation}\label{eq:BasisFunctionSeparation}
	\Delta_{\phi} \leq \frac{1}{2\rho_{\phi}\boldsymbol{\nu}_{c}}
\end{equation}
where $\rho_{\phi} \in \mathbb{R} \ge 1$ is an oversampling parameter to determine the basis function separation.

\subsection{Estimating Support For Connectivity Kernel}
The support for the connectivity kernel can be inferred by studying the cross-correlation between consecutive observations of the neural field. 
\begin{itemize}
	\item it is difficult to tell what correlations are from the cross-talk due the the sensor kernel
	\item it is difficult because of the nonlinear function
	\item 
\end{itemize}

@@

\Xi_2&=\sum_{t=1}^T\Lambda_t(\mathbf r')\mathbf P_t \tilde{\Lambda}_t(\mathbf r') \\

@@

\section{Discussion}
\begin{itemize}
	\item  
\end{itemize}

>>>>

\title{Patient-Specific Neural Field Modeling from Micro-Electrode Data}
\author{In no specific order: Parham Aram, Dean R. Freestone, Kenneth Scerri, Michael Dewar,\\
 Andrew Zammit Mangion, Jacob A. Donoghue, Sydney S. Cash, Leigh R. Hochberg,\\
 David B. Grayden and Visakan Kadirkamanathan  }

@@

During the time-course of the iEEG monitoring period, several seizures were recorded. This study focuses on the local field potential (LFP) data recorded with the micro-electrode during a seizure.

The first pre-processing step was to identify channels that had very poor signal quality. These channels were set to the $nan$ data type. Following this, the data was re-referenced to a common average montage to remove common mode artifact and the effect of the reference electrode. Next the data was low-pass filtered with a cut-off frequency of 100~Hz. The data was then resampled from 30~kHz to 5~kHz. To enable spatial visualisation and spatial frequency analysis, the corrupted channels that were set to $nan$, were interpolated using data from all neighbouring channels.

@@

The spectral properties of the recorded time series for each of the time segments are shown in Figure~\ref{fig:TemporalFreqObservation}. Each plot shows the averaged (across channels) spectrum of the data. The pre-seizure period shows a decay in power at a typical $1/f^2$ rate. The seizure period has higher power at low and high ends of the spectrum. In addition, there is a peak in the power of the LFPs at ~51~Hz during the seizure. This peak would be difficult to observe in European, Asian and other countries due to the 50~Hz mains power frequency. The post-seizure period shows a large increase in lower frequency power...

Figure~\ref{fig:SpatialFreqObservation} shows the results of spatial frequency analysis of the observed neural field. Throughout this paper the spatial frequency shall be denoted by $\nu$. Each subplot is the average (across time) spatial frequency during either the pre-seizure, seizure or post-seizure periods. The spatial frequency of the neural field is governed by disturbances (inputs from outside the field) and the underlying connectivity structure. The symmetry in the plots provides some evidence for homogeneous and isotropic connectivity.

Figure~\ref{fig:DiagSpatialFreqObservation} shows diagonal cross-sections of the spatial frequency plots. From this figure we can see that the spatial cut-off frequency (-3~dB point) is $\nu_c \approx 0.26$~Hz. The profile of the diagonal cross-section indicates that the field is adequately sampled from the micro-electrode array, where the power drops by approximately 25~dB at the maximum observable spatial frequency (according to Shannon's sampling theorem) indicating spatial aliasing will have an insignificant effect on the sampled data. The spatial cut-off frequency provides information on the complexity of the model required to represent the neural field. Considering the field is band-limited, we can represent it by a finite set of continuous basis functions. According the Shannon's sampling theorem, the minimum spacing between basis functions, $\Delta_{\phi}$, to describe the field is
\begin{equation}\label{eq:BasisFunctionSeparation}
	\Delta_{\phi} \leq \frac{1}{2\rho_{\phi}\boldsymbol{\nu}_{c}}
\end{equation}
where $\rho_{\phi} \in \mathbb{R} \ge 1$ is an oversampling parameter to determine the basis function separation.

\subsection{Estimating Support For Connectivity Kernel}
Under the assumption that the sensors are not spatially band-limiting the spectral content of the field, the support for the connectivity kernel can be inferred by studying the spatial cross-correlation between consecutive observations of the neural field. The deterministic component of the spatial mapping between consecutive fields is determined by the connectivity kernel convolved with the firing rates. Although the relationship between the field and the firing rate is nonlinear, the cross-correlation analysis still extracts meaningful results. This can be seen by considering the three operating regions of the sigmoid. When the sigmoid is mean membrane voltage is below the active region of the sigmoid, the firing rate will be approximately zero and only very small spatial correlations will observed. Therefore, this will not contribute to the results. Within the active region the sigmoid can be approximated by a linear function between consecutive samples, given a sufficiently high sampling rate. Therefore, linear cross-correlation analysis is appropriate. When the sigmoid is in the saturated region, the correlation coefficients will be higher. Therefore, we can not estimate the gain for the kernel, but the spatial extent can still be inferred. The gain for the kernel will be inferred using the EM algorithm described below.

Figure~\ref{fig:SpatialCrossCorrelation} shows the average (across time) cross-correlations between consecutive field observations for the different time periods relative to the seizure progression. By averaging over time, the effect of the disturbance on the analysis is reduced. The figure shows that the connectivity kernel has a local excitation with surround inhibition structure. can be approximated by a local as a two-dimensional Gaussian. The wi
\begin{itemize}
	\item it is difficult to tell what correlations are from the cross-talk due the the sensor kernel
	\item it is difficult because of the nonlinear function
	\item 
\end{itemize}

@@

\Xi_2&=\sum_{t=1}^T\left[q^\top(\mathbf{\hat x}_t)\boldsymbol\Sigma_e^{-1}q(\mathbf{\hat x}_t)+\Lambda_t(\mathbf r')\mathbf P_t \tilde{\Lambda}_t(\mathbf r')\right] \\

@@

\section{Discussion}

\begin{itemize}
	\item  
\end{itemize}
